### 2019-12-27 Statistical Methods/08_Regression.rmd
statistical methods, ch8
statistical methods

Assume normality and homoscedasticity and suppose $n = 38, \bar{Y} = 20, \sum{X^2_i} = 1,922, \bar{X} = 7, \sum{(X_i - \bar{X})(Y_i - \bar{Y})} = 180, \sum{X_i - \bar{X}^2} = 60, s^2_{X, Y} = 121.$

@@

```{r, echo = T}
```

b.) Test the hypothesis $H_0 : \beta_0 = 0, \alpha = 0.05$

```{r, echo = T}
```

@@

```{r, echo = T}
```

@@

@@

@@

@@

@@

@@

@@

@@

@@

@@

@@

@@

### 8.31
### 8.32
### 8.33
### 8.34
### 8.35
### 8.36

>>>>

Assume normality and homoscedasticity and suppose $n = 38, \bar{Y} = 20, \sum{X^2_i} = 1,922, \bar{X} = 7, \sum{(X_i - \bar{X})(Y_i - \bar{Y})} = 180, \sum{(X_i - \bar{X})^2} = 60, s^2_{X.Y} = 121.$

@@

```{r, echo = T}
n <- 38; ybar <- 20; xbar <- 7
variance <- 121; A <- 180; C <- 60
ssx <- 1922

b.) Test the hypothesis $H_0 : \beta_0 = 0, \alpha = 0.02$

```{r, echo = T}
alpha <- 0.02

slope + qt(c(Lower = alpha/2, Upper = 1 - alpha/2), df = n - 2) * sqrt(variance / n)

Tval <- intercept * sqrt( ((n - 2) * C) / (variance  * ssx) )

crit <- qt(alpha/2, df = n - 2)

ifelse(abs(Tval) < crit, "Reject Null", "Fail to Reject")
```

@@

```{r, echo = T}
alpha <- 1 - .9

slope + qt(c(Lower = alpha/2, Upper = 1 - alpha/2), df = n - 2) * sqrt( variance / C)
```

@@

```{r, echo = T}
n <- 41; ybar <- 10; xbar <- 12; 
ssr <- 100; ssx <- 400; variance <- 144

b1 <- ssr / ssx

b0 <- ybar - xbar*b1
```

$\beta_0 = `r b0`, \beta_1 = `r b1`$

@@

```{r, echo = T}
alpha <- 1 - .9

b1 + qt(c( Lower = alpha/2, Upper = 1 - alpha/2), df = n - 2) * sqrt( variance / ssx)
```

@@

```{r, echo = T}
n <- 18; b1 <- 3.1; ssx <- 144; variance <- 36
alpha <- 1 - .95

b1 + qt(c(Lower = alpha/2, Upper = 1 - alpha/2), df = n - 2) * sqrt( variance / ssx )

@@

_Given the confidence interval above, it seems reasonable._

@@

```{r, echo = T}
n <- 20; b0 <- 6; ssx <- 169
variance <- 25; xvar <- 90
alpha <- 1 - .95

b0 + qt(c(Lower = alpha/2, Upper = 1 - alpha/2), df = n - 2) * sqrt( (variance*ssx) / (n * xvar))
```

@@

```{r, echo = T}
n <- 27; yvar <- 100; xvar <- 625
ssr <- 200; alpha <- 0.01

r <- ssr / sqrt( yvar * xvar )

test.stat.T <- r * sqrt( (n-2)/ (1 - r^2))

crit <- qt(1 - alpha/2, df = n - 2)

ifelse(abs(test.stat.T) >= crit, "Reject Null", "Fail to Reject")
```

@@

```{r, echo = T}
n <- 5; yvar <- 16; xvar <- 25
ssr <- 10; alpha <- 0.05

r <- ssr / sqrt( yvar * xvar )

test.stat.T <- r * sqrt( (n-2)/ (1 - r^2))

crit <- qt(1 - alpha/2, df = n - 2)

ifelse(abs(test.stat.T) >= crit, "Reject Null", "Fail to Reject")
```

@@

The high school grade-point average (X) and college grade-point (Y) for 29 randomly sampled college freshmen yielded the following results:

$\sum{(Y_i - \bar{Y})}^2 = 64, \sum{(X_i - \bar{X})}^2 = 100, \sum{(X_i - \bar{X})(Y_i - \bar{Y})} = 40$

Test $H_0 : \rho = 0, \alpha = 0.1$ and interprest the results.

```{r, echo = T}
n <- 29; yvar <- 64; xvar <- 100; ssr <- 40
alpha <- 0.1

r <- ssr / sqrt( yvar * xvar )

crit <- qt(1 - alpha/2, df = n - 2)

test.stat.T <- r * sqrt( (n - 2) / (1 - r^2) )

ifelse( abs(test.stat.T) >= crit, "Reject null", "Fail to reject")
```

@@

For the previous exercise, answer the following questions:

a.) Is it reasonable to conclude that the least squares regression line has a positive slope?

_Yes, it appears there is a positive correlation._

b.) Is it possible that despite the value for r, as high school grade-point average increases, college grade-point average decreases? Explain.

_It is possible, although unlikely._

c.) What might you do, beyond considering r, to decide it is reasonable to conclude that as high school grade-point averages increase, college grade point averages increase as well?

_Plot the data to assess the visual indication of the relationship._

@@

Using R, determine what happens to Pearson's correlation between X and Y if the Y values are multiplied by 3. Argue that if Y is multiplied by any constant c != 0, Pearson's correlation does not change.

```{r, echo = T}
set.seed(46)
x <- rnorm(30); y <- rnorm(30)
cor(x, y)

cor(3*x, y)

@@

Repeat the previous problem, only determine what happens to the slope of the least squares regresion line.

```{r, echo = T}
set.seed(46)
x <- rnorm(30); y <- rnorm(30)

lsfit(x, y)$coef

lsfit(x, 3*y)$coef
```

The absolute value of the slope gets larger.

@@

Consider a least squares regression line $Y = 0.52X + 2$, assume homoscedasiticy as consider the situation where the common variance is $\sigma^2 = 1$? What happens to the correlation coefficent between X and Y if instead $\sigma^2 = 2$?

_It will increase the variance so the correlation will decrease._

### 8.31

The numerator of the coefficient of determination is $\sum{(Y_i - \bar{Y}) - \sum{(Y_i - \hat{Y})}}$. Based on the least squares principal, why is the value always greater than zero?

_If $\bar{Y}$ is used to predict Y, $\sum{(Y_i - \bar{Y})}$ will be larger than $\sum{(Y_i - \hat{Y})}$._

### 8.32

Imagine a study where the correlation between some amount of an experimental drug and liver damage yields a value for r close to zero and the hypothesis $H_0 : \rho = 0$ is not rejected. Why might it be unreasonable to conclude that the two variabels are independent?

_The magnitude of the residuals, curvature, outliers._

### 8.33

Suppose $r^2 = 0.95$. Explain why this does not provide convincing evidence that the least squares line provides a good fit to a scatterplot of the points.

_Outliers can cause large $r^2$, however, be a poor fit._

### 8.34

Imagine a situation where points are removed for which the X values are judged to be outliers. Note that this restricts the range of X values. Without looking at the data, can someone predict whether Pearson's correlation will increase or decrease after these points are removed?

_No. You need to look at the data. Restricting the range of X can increase as well as decrease r._

### 8.35

If the normality assumption is violated, what effect might this have when computing confidence intervals for the slope and intercept as described in 8.4.1?

_The condence interval can be relatively long and is potentially inaccurate._

### 8.36

If the homoscedasticity assumption is violated, what effect might this have when computing confidence intervals as described in 8.4.1?

_The condence interval can be relatively inaccurate due to using the wrong standard error._
